---
title: "creative-data-visualization"
author: "Drew Hynes"
format: html
editor: visual
---

## Data Wrangling

```{r}
library(readr)
library(dplyr)
library(skimr)

#read in dataframes
m <- "UT27_Meninges_Output_Data.csv"
meninges <- read_csv(m, col_names = TRUE)
b <- "UT27_Brain_Output_Data.csv"
brain <- read_csv(b,col_names = TRUE)

#keep only needed columns
brain_select <- brain %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Microglia Counts`, `Neutrophil Counts`, `T-Cell Counts`, `CD8+ T Cell Counts`, `CD4+ T Cell Counts`, `B Cell Counts`)
meninges_select <- meninges %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Neutrophil Cells`, `T Cell Cells`, `CD8 Cells`, `CD4 Cells`, `B Cell Cells`)

#Make column names consistent
meninges_select <- meninges_select %>%
  rename(
    `Meninges Neutrophil Counts` = `Neutrophil Cells`,
    `Meninges T-Cell Counts` = `T Cell Cells`,
    `Meninges CD8+ T Cell Counts` = `CD8 Cells`,
    `Meninges CD4+ T Cell Counts` = `CD4 Cells`,
    `Meninges B Cell Counts` = `B Cell Cells`
  )
brain_select <- brain_select %>%
  rename(
    `Brain Microglia Counts` = `Microglia Counts`,
    `Brain Neutrophil Counts` = `Neutrophil Counts`,
    `Brain T-Cell Counts` = `T-Cell Counts`,
    `Brain CD8+ T Cell Counts` = `CD8+ T Cell Counts`,
    `Brain CD4+ T Cell Counts` = `CD4+ T Cell Counts`,
    `Brain B Cell Counts` = `B Cell Counts`
  )

#combine cage and mouse ID into 1 column
brain_select$`Mouse #` <- paste(brain_select$`Cage #`, brain_select$`Mouse #`, sep = "-")
meninges_select$`Mouse #` <- paste(meninges_select$`Cage #`, meninges_select$`Mouse #`, sep = "-")

#join the 2 datasets
combined_data <- brain_select %>%
  full_join(meninges_select) 

#Narrow the data to just groups of interest
combined_data <- na.omit(combined_data)
combined_data <- combined_data %>%
  filter(!(Group %in% c("2-WT-9", "5-3xTg-18")))

#PT2

library(scales)

#Combine subjects of same group to make one group average amount of each immune cell type
averaged_data <- combined_data %>%
  group_by(Group, ZT) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE), .groups = "drop")

#Going to change the counts to a scaled number 0.1-1.0 that represents the data
min_size <- 0.1   # smallest size for visualization
max_size <- 1.0  # largest size for visualization

# List of columns to rescale
cell_count_columns <- c(
  "Meninges Neutrophil Counts",
  "Meninges T-Cell Counts",
  "Meninges CD8+ T Cell Counts",
  "Meninges CD4+ T Cell Counts",
  "Meninges B Cell Counts",
  "Brain Microglia Counts",
  "Brain Neutrophil Counts",
  "Brain T-Cell Counts",
  "Brain CD4+ T Cell Counts",
  "Brain B Cell Counts",
  "Brain CD8+ T Cell Counts"
)

# Rescale specified columns
averaged_data <- averaged_data %>%
  mutate(across(all_of(cell_count_columns), ~ rescale(., to = c(min_size, max_size))))

#PT3

library(dplyr)
library(tidyr)

# Add cell type → icon mapping
icon_paths <- tibble(
  cell_type = c(
    "Meninges Neutrophil Counts",
    "Meninges T-Cell Counts",
    "Meninges CD8+ T Cell Counts",
    "Meninges CD4+ T Cell Counts",
    "Meninges B Cell Counts",
    "Brain Microglia Counts",
    "Brain Neutrophil Counts",
    "Brain T-Cell Counts",
    "Brain CD4+ T Cell Counts",
    "Brain B Cell Counts",
    "Brain CD8+ T Cell Counts"
  ),
  icon_path = c(
    "neutrophil.png",
    "tcell.png",
    "cd8tcell.png",
    "cd4tcell.png",
    "bcell.png",
    "microglia.png",
    "neutrophil.png",
    "tcell.png",
    "cd4tcell.png",
    "bcell.png",
    "cd8tcell.png"
  )
)

# Pivot the averaged data to long format to map icons
long_averaged <- averaged_data %>%
  pivot_longer(cols = all_of(cell_count_columns),
               names_to = "cell_type",
               values_to = "scaled_size") %>%
  left_join(icon_paths, by = "cell_type")


library(dplyr)
library(stringr)

#create column for tissue type so that they can be split later
long_averaged <- long_averaged %>%
  mutate(tissue_type = case_when(
    str_detect(cell_type, "Brain") ~ "Brain",
    str_detect(cell_type, "Meninges") ~ "Meninges",
    TRUE ~ "Other"
  ))

long_averaged <- long_averaged %>%
  mutate(scaled_size = round(scaled_size, 1))  # Round to 1 decimal place (tenths) to help with visualization

# Split the dataframe into two based on tissue_type
brain_averaged <- long_averaged %>% 
  filter(tissue_type == "Brain")
meninges_averaged <- long_averaged %>% 
  filter(tissue_type == "Meninges")

#split each of these into dataframes for each group
# For brain_averaged
brain_WT6      <- brain_averaged %>% filter(Group == "1-WT-6")
brain_3xTg9    <- brain_averaged %>% filter(Group == "3-3xTg-9")
brain_WT18     <- brain_averaged %>% filter(Group == "4-WT-18")

# For meninges_averaged
meninges_WT6   <- meninges_averaged %>% filter(Group == "1-WT-6")
meninges_3xTg9 <- meninges_averaged %>% filter(Group == "3-3xTg-9")
meninges_WT18  <- meninges_averaged %>% filter(Group == "4-WT-18")


```

Split into dataframe based on cell type

```{r}
# Define the target icons
icons <- c("microglia.png", "tcell.png", "cd8tcell.png", "cd4tcell.png", "bcell.png", "neutrophil.png")

# ---- For brain dataframes ----
brain_WT6_microglia     <- brain_WT6      %>% filter(icon_path == "microglia.png")
brain_WT6_tcell         <- brain_WT6      %>% filter(icon_path == "tcell.png")
brain_WT6_cd8tcell      <- brain_WT6      %>% filter(icon_path == "cd8tcell.png")
brain_WT6_cd4tcell      <- brain_WT6      %>% filter(icon_path == "cd4tcell.png")
brain_WT6_bcell         <- brain_WT6      %>% filter(icon_path == "bcell.png")
brain_WT6_neutrophil    <- brain_WT6      %>% filter(icon_path == "neutrophil.png")

brain_3xTg9_microglia   <- brain_3xTg9    %>% filter(icon_path == "microglia.png")
brain_3xTg9_tcell       <- brain_3xTg9    %>% filter(icon_path == "tcell.png")
brain_3xTg9_cd8tcell    <- brain_3xTg9    %>% filter(icon_path == "cd8tcell.png")
brain_3xTg9_cd4tcell    <- brain_3xTg9    %>% filter(icon_path == "cd4tcell.png")
brain_3xTg9_bcell       <- brain_3xTg9    %>% filter(icon_path == "bcell.png")
brain_3xTg9_neutrophil  <- brain_3xTg9    %>% filter(icon_path == "neutrophil.png")

brain_WT18_microglia    <- brain_WT18     %>% filter(icon_path == "microglia.png")
brain_WT18_tcell        <- brain_WT18     %>% filter(icon_path == "tcell.png")
brain_WT18_cd8tcell     <- brain_WT18     %>% filter(icon_path == "cd8tcell.png")
brain_WT18_cd4tcell     <- brain_WT18     %>% filter(icon_path == "cd4tcell.png")
brain_WT18_bcell        <- brain_WT18     %>% filter(icon_path == "bcell.png")
brain_WT18_neutrophil   <- brain_WT18     %>% filter(icon_path == "neutrophil.png")

# ---- For meninges dataframes  ----
meninges_WT6_tcell         <- meninges_WT6     %>% filter(icon_path == "tcell.png")
meninges_WT6_cd8tcell      <- meninges_WT6     %>% filter(icon_path == "cd8tcell.png")
meninges_WT6_cd4tcell      <- meninges_WT6     %>% filter(icon_path == "cd4tcell.png")
meninges_WT6_bcell         <- meninges_WT6     %>% filter(icon_path == "bcell.png")
meninges_WT6_neutrophil    <- meninges_WT6     %>% filter(icon_path == "neutrophil.png")

meninges_3xTg9_tcell       <- meninges_3xTg9   %>% filter(icon_path == "tcell.png")
meninges_3xTg9_cd8tcell    <- meninges_3xTg9   %>% filter(icon_path == "cd8tcell.png")
meninges_3xTg9_cd4tcell    <- meninges_3xTg9   %>% filter(icon_path == "cd4tcell.png")
meninges_3xTg9_bcell       <- meninges_3xTg9   %>% filter(icon_path == "bcell.png")
meninges_3xTg9_neutrophil  <- meninges_3xTg9   %>% filter(icon_path == "neutrophil.png")

meninges_WT18_tcell        <- meninges_WT18    %>% filter(icon_path == "tcell.png")
meninges_WT18_cd8tcell     <- meninges_WT18    %>% filter(icon_path == "cd8tcell.png")
meninges_WT18_cd4tcell     <- meninges_WT18    %>% filter(icon_path == "cd4tcell.png")
meninges_WT18_bcell        <- meninges_WT18    %>% filter(icon_path == "bcell.png")
meninges_WT18_neutrophil   <- meninges_WT18    %>% filter(icon_path == "neutrophil.png")

```

Function to plot prevalence of Immune cell represented by size, chronologically left to right. Can do this for each of the above dataframes by inputting its name and the icon name

```{r}
library(ggplot2)
library(ggimage)
library(magick)

# Function to plot icons with scaling and spacing
plot_icons <- function(dataframe, icon_image) {
  
  # Create the df_icons data frame based on the dataframe input
  df_icons <- data.frame(
    x = c(1, 2, 3, 4),   
    y = c(1, 1, 1, 1),  
    image = rep(icon_image, nrow(dataframe)),  # Use icon_image input
    scaled_size = dataframe$scaled_size  # Reference the 'scaled_size' column from the input dataframe
  )
  
  # Plot icons left to right chronologically with scaled size
  p <- ggplot(df_icons, aes(x = x, y = y)) +
    geom_image(aes(image = image, size = scaled_size), by = "height") +  # Use actual sizes from scaled_size column
    scale_size_identity() +  # Use actual sizes from scaled_size column
    theme_void() +  # Remove grid 
    theme(
      axis.text = element_blank(),  
      axis.ticks = element_blank()  
    ) +
    coord_cartesian(xlim = c(0.5, 5.5), ylim = c(0.5, 1.5))  # Fixed xlim for a consistent width, so you can stack this and compare to other plots
  
  # Print the plot
  print(p)
}

```

Example Usage

```{r}
library(ggplot2)     
library(ggimage)     
library(patchwork)    


# Example usage:
# Assuming meninges_WT6_tcell is your dataframe and "tcell.png" is the image
meninges_WT6_tcell_plot <- plot_icons(meninges_WT6_tcell, "tcell.png")

meninges_3xTg9_tcell_plot <- plot_icons(meninges_3xTg9_tcell, "tcell.png")

meninges_WT18_tcell_plot <- plot_icons(meninges_WT18_tcell, "tcell.png")

# Add titles/labels for each plot to specify the group
meninges_WT6_tcell_plot <- meninges_WT6_tcell_plot + ggtitle("WT6 Group")
meninges_3xTg9_tcell_plot <- meninges_3xTg9_tcell_plot + ggtitle("3xTg9 Group")
meninges_WT18_tcell_plot <- meninges_WT18_tcell_plot + ggtitle("WT18 Group")

# Stack the plots vertically with patchwork
combined_meninges_tcell_plot <- (meninges_WT6_tcell_plot / meninges_3xTg9_tcell_plot / meninges_WT18_tcell_plot)

print(combined_meninges_tcell_plot)

#This plot shows the amount of t cells in the meninges, scaled by size of the T cell image, over the course of a day. The left side is the beginning of the day, and the right is the end of the day. This allows me to compare the amount of T cells throughout the day between adult mice, alzheimer's model mice, and aged mice, to see when they may be most disrupted in the models
```

Another example, for brain CD8 T cells

```{r}
library(ggplot2)     
library(ggimage)     
library(patchwork)    

# Example usage:
# Assuming brain_WT6_cd8tcell, brain_3xTg9_cd8tcell, brain_WT18_cd8tcell are your dataframes and "cd8tcell.png" is the image
brain_WT6_cd8tcell_plot <- plot_icons(brain_WT6_cd8tcell, "cd8tcell.png")
brain_3xTg9_cd8tcell_plot <- plot_icons(brain_3xTg9_cd8tcell, "cd8tcell.png")
brain_WT18_cd8tcell_plot <- plot_icons(brain_WT18_cd8tcell, "cd8tcell.png")

# Add titles/labels for each plot to specify the group
brain_WT6_cd8tcell_plot <- brain_WT6_cd8tcell_plot + ggtitle("WT6 Group")
brain_3xTg9_cd8tcell_plot <- brain_3xTg9_cd8tcell_plot + ggtitle("3xTg9 Group")
brain_WT18_cd8tcell_plot <- brain_WT18_cd8tcell_plot + ggtitle("WT18 Group")

# Stack the plots vertically with patchwork
combined_brain_cd8tcell_plot <- (brain_WT6_cd8tcell_plot / brain_3xTg9_cd8tcell_plot / brain_WT18_cd8tcell_plot)

# Print the final combined plot
print(combined_brain_cd8tcell_plot)
#This does the same, but for brain CD8+ T cells
```

This github has image files and all data needed to do this for all of the different cell types as needed
